
<font size="4">


## Getting recordings from Xeno-Canto


The warbleR function [`query_xc()`](https://marce10.github.io/warbleR/reference/query_xc.html) queries for avian vocalization recordings in the open-access online repository [Xeno-Canto](https://xeno-canto.org). It can return recordings metadata or download the associated sound files. 

Example on how to get recording's metadata (no downloading):

```{r, message=FALSE,  warning=FALSE}

library(warbleR)

pl <- query_xc(qword = 'Parus major', download = FALSE)

```

&nbsp;

Keep only song vocalizations: 
```{r}

song_pl <- pl[grep("song", ignore.case = TRUE, pl$Vocalization_type), ]

```

&nbsp;

Map locations using [`map_xc()`](https://marce10.github.io/warbleR/reference/map_xc.html):
```{r, results = 'asis', fig.width=10}

map_xc(song_pl, leaflet.map = TRUE)

```
&nbsp;

Once you feel fine with the subset of data you can go ahead and download it as follows:
```{r, eval = FALSE, fig.width=10}

query_xc(X = song_pl, download = TRUE, path = "DIRECTORY WHERE YOU WANT TO SAVE THE FILES")

```
&nbsp;


<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```
