---
date: "`r Sys.Date()`"
toc: true
toc-depth: 3
toc-location: left
number-sections: true
highlight-style: pygments
format:
  html:
    df-print: kable
    code-fold: show
    code-tools: true
    css: styles.css
    link-external-icon: true
    link-external-newwindow: true      
---

 

::: {.alert .alert-info}
## **Objetive** {.unnumbered .unlisted}

-   Demonstrate how to obtain acoustic data from online repositories
:::

 

<font size="4">

## Getting recordings from Xeno-Canto

The warbleR function [`query_xc()`](https://marce10.github.io/warbleR/reference/query_xc.html) queries for avian vocalization recordings in the open-access online repository [Xeno-Canto](https://xeno-canto.org). It can return recordings metadata or download the associated sound files.

Example on how to get recording's metadata (no downloading):

```{r, message=FALSE,  warning=FALSE}

library(warbleR)

pl <- query_xc(qword = 'Parus major', download = FALSE)

```

Keep only song vocalizations:

```{r}

song_pl <- pl[grep("song", ignore.case = TRUE, pl$Vocalization_type), ]

```

Map locations using [`map_xc()`](https://marce10.github.io/warbleR/reference/map_xc.html):

```{r, results = 'asis', fig.width=10}

map_xc(song_pl, leaflet.map = TRUE)

```

Once you feel fine with the subset of data you can go ahead and download it as follows:

```{r, eval = FALSE, fig.width=10}

query_xc(X = song_pl, download = TRUE, path = "DIRECTORY WHERE YOU WANT TO SAVE THE FILES")

```

------------------------------------------------------------------------

<font size="4">Session information</font>

```{r session info, echo=F}

sessionInfo()

```
